---
title: "DATA 607 Week 1 Assignment – Loading Data into a Data Frame"
author: "Beshkia Kvarnstrom"
date: "2023-01-29"
output:
  html_document: default
  pdf_document: default
---


```{r DATA607-Assignment1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(plyr)
```

 Task - Choose one of the provided datasets on fivethirtyeight.com that you find interesting: https://data.fivethirtyeight.com/
 
 You should first study the data and any other information on the GitHub site, and read the associated
 fivethirtyeight.com article.
 
 To receive full credit, you should:
 1. Take the data, and create one or more code blocks. You should finish with a data frame that contains a
    subset of the columns in your selected dataset. If there is an obvious target (aka predictor or independent)
    variable, you should include this in your set of columns. You should include (or add if necessary)
    meaningful column names and replace (if necessary) any non-intuitive abbreviations used in the data that
    you selected. For example, if you had instead been tasked with working with the UCI mushroom dataset,
    you would include the target column for edible or poisonous, and transform “e” values to “edible.” Your
    deliverable is the R code to perform these transformation tasks.
 
 2. Make sure that the original data file is accessible through your code—for example, stored in a GitHub
    repository or AWS S3 bucket and referenced in your code. If the code references data on your local
    machine, then your work is not reproducible!
 
 3. Start your R Markdown document with a two to three sentence “Overview” or “Introduction” description of
    what the article that you chose is about, and include a link to the article.
 
 4. Finish with a “Conclusions” or “Findings and Recommendations” text block that includes what you might do
    to extend, verify, or update the work from the selected article.
 
 5. Each of your text blocks should minimally include at least one header, and additional non-header text.

 6. You’re of course welcome—but not required--to include additional information, such as exploratory data
    analysis graphics (which we will cover later in the course).

 7. Place your solution into a single R Markdown (.Rmd) file and publish your solution out to rpubs.com.

 8. Post the .Rmd file in your GitHub repository, and provide the appropriate URLs to your GitHub repositor y
    and your rpubs.com file in your assignment link.
 
 OVERVIEW
 The Data I will be working with for this assignment compare the comsuption of wine, beer and other spirits of different countries across the   world. A row is listed for each country along with the amount of each alcoholic beverage.

 The dataset was taken from: https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits// 
 
```{r Load Data, echo=TRUE}

## Load data from GitHub

AlcoholConsumptionData = read.table(file="https://raw.githubusercontent.com/BeshkiaKvarnstrom/MSDS-DATA607/main/drinks.csv", header=TRUE,sep=",")
```


```{r Read Data from DF, echo=TRUE}
# Reading the data in the Dataframe
AlcoholConsum <- data.frame(AlcoholConsumptionData$country, AlcoholConsumptionData$beer_servings, AlcoholConsumptionData$spirit_servings, AlcoholConsumptionData$wine_servings, AlcoholConsumptionData$total_litres_of_pure_alcohol)

colnames(AlcoholConsum) <- c("COUNTRY", "BEER", "SPIRIT", "WINE", "TOTAL LITRES")

AlcoholConsum
```


```{r echo=TRUE}
# Countries without alcohol consumption
NoConsumtion_sub <- subset(AlcoholConsum, BEER == 0.0 & WINE == 0.0)
NoConsumtion_sub <- NoConsumtion_sub[, c("COUNTRY", "BEER", "WINE")]
NoConsumtion_sub
```

 
```{r echo=TRUE}
# Countries that consume alcohol

AlcoholConsum_sub <- subset(AlcoholConsum, BEER > 0 & WINE > 0 )
AlcoholConsum_sub <- AlcoholConsum_sub[, c("COUNTRY", "BEER", "WINE")]
AlcoholConsum_sub
```

```{r}
# Top 10 countries with highest consumption of alcohol pe liter

TopConsum <- head(arrange(AlcoholConsum, desc(AlcoholConsum$`TOTAL LITRES`)), n = 10)
TopConsum <- TopConsum[, c("COUNTRY", "BEER", "WINE")]
TopConsum
```



```{r}
Consumgg <- ggplot(NULL, aes(x, y)) + 
geom_line(data=TopConsum, aes(x=COUNTRY, y=WINE, group=1), col="green") +
geom_line(data=TopConsum, aes(x=COUNTRY, y=BEER, group=1), col="purple")+
  labs(title = "WINE & BEER CONSUMPTION BY COUNTRY", x= "COUNTRY", y = "TOTAL SERVINGS", ylim=c(100,1000), breaks = 10) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# display the plot
Consumgg
```


CONCLUSION

The study shows that based on the top 10 total consumption by litre, beer was more popular than wine. Out Of the 194 countries 14 did not consume any alcohol. 
